### 2016-05-25 cv.tex
Updates for the review-ready version

\cventry{2011--Nu}{Senior Software Ontwikkelaar}{SDL Fredhopper}{Amsterdam}{}{Groep leider in cloud\newline}

>>>>

\cventry{mei 2016}{Lead Software Ontwikkelaar}{HERE}{Eindhoven}{}{CTO Office Applied Technologies\newline}

\cventry{2011--apr. 2016}{Senior Software Ontwikkelaar}{SDL Fredhopper}{Amsterdam}{}{Groep leider in cloud\newline}

### 2016-05-25 p00-ch00-intro/intro.tex
Update abstract based on comments from Stijn
Updates for the review-ready version

In a concurrent setting with objects, multi-threading has been a common approach to provide an object-oriented concurrency model.
A thread holds a single stack of synchronous method calls.
Execution in a thread is sequential.
The stack has a starting location (i.e. start of invocation) and a finish location (i.e. the return point).
Future values can be used to hold the eventual return value of a method at the call site~\cite{DBLP:books/daglib/0092230,de2007complete}.
In a caller-callee setting, an object calls methods from other objects.
If multiple threads execute simultaneously in an object, their instructions are
interleaved in an uncontrolled manner.
In coroutines, the points of interleaving are explicit (i.e. \emph{yield}) 
whereas multi-threading is based on implicit scheduling of method invocations.
Furthermore, interactions in both coroutines and multi-threading are blocking and synchronous.
In contrast, the actor model relies on asynchronous communication (which is non-blocking)
as one of its core elements.

@@

The core contributions of this thesis target the intersection of object 
orientation, actor model, and concurrency.
We choose Java as the main programming language and as one of the mainstream 
object-oriented languages. 
We formalize a subset of Java and its concurrency API~\cite{jsr166} to 
facilitate formal verification and reasoning about it (cf. Chapter~\ref{ch:p01:ch02}).
We create an abstract mapping from a concurrent-object modeling language, 
ABS~\cite{johnsen2012abs}, to the programming semantics of concurrent Java (cf. Chapter~\ref{ch:p01:ch02}). 
We provide the formal semantics of the mapping and runtime properties of 
the concurrency layer including deadlines and scheduling policies (cf. Chapter~\ref{ch:p01:ch01}).
We provide an implementation of the ABS concurrency layer as a Java API library 
and framework utilizing the latest language additions 
in Java 8~\cite{jsr335:lambda:translation} (cf. Chapter~\ref{ch:p02:ch01}).
We design and implement a runtime monitoring framework, JMSeq, to verify the
correct ordering of execution of methods through code annotations in JVM (cf. Chapter~\ref{ch:p03:ch02}). 
In addition, we design a large-scale monitoring system as a real-world 
application; the monitoring system is built with ABS concurrent objects 
and formal semantics that leverages schedulability 
analysis to verify correctness of the monitors~\cite{fersman2007task} (cf. Chapter~\ref{ch:p03:ch01}).
Table~\ref{tbl:thesis} summarizes the structure of this text.

>>>>

In a concurrent setting with objects, multi-threading is a common approach to provide an object-oriented concurrency model.
A thread holds a single stack of synchronous method calls.
Execution in a thread is sequential.
The stack has a starting location (i.e. start of invocation) and a finish location (i.e. the return point).
Future values can be used to hold the eventual return value of a method at the call site~\cite{DBLP:books/daglib/0092230,de2007complete}.
In a caller-callee setting, an object calls methods from other objects.
If multiple threads execute simultaneously in an object, their instructions are
interleaved in an uncontrolled manner.
In coroutines, the points of interleaving are explicit (i.e. \emph{yield}) 
whereas multi-threading is based on implicit scheduling of method invocations.
Furthermore, interactions in both coroutines and multi-threading are blocking and synchronous.
In contrast, the actor model relies on asynchronous communication (which is non-blocking)
as one of its core elements.

@@

The core contributions of this thesis target the intersection of object 
orientation, actor model, and concurrency.
We choose Java as the main target programming language and as one of the mainstream 
object-oriented languages. 
We formalize a subset of Java and its concurrency API~\cite{jsr166} to 
facilitate formal verification and reasoning about it (cf. Chapter~\ref{ch:p01:ch02}).
We create an abstract mapping from a concurrent-object modeling language, 
ABS~\cite{johnsen2012abs}, to the programming semantics of concurrent Java (cf. Chapter~\ref{ch:p01:ch02}). 
We provide the formal semantics of the mapping and runtime properties of 
the concurrency layer including deadlines and scheduling policies (cf. Chapter~\ref{ch:p01:ch01}).
We provide an implementation of the ABS concurrency layer as a Java API library 
and framework utilizing the latest language additions 
in Java 8~\cite{jsr335:lambda:translation} (cf. Chapter~\ref{ch:p02:ch01}).
We design and implement a runtime monitoring framework, JMSeq, to verify the
correct ordering of execution of methods through code annotations in JVM (cf. Chapter~\ref{ch:p03:ch02}). 
In addition, we design a large-scale monitoring system as a real-world 
application; the monitoring system is built with ABS concurrent objects 
and formal semantics that leverages schedulability 
analysis to verify correctness of the monitors~\cite{fersman2007task} (cf. Chapter~\ref{ch:p03:ch01}).
Table~\ref{tbl:thesis} summarizes the structure of this text.
